Acceleration of Derivative Calculations with Application to Radial Basis Function ­ Finite-Differences on the Intel MIC Architecture

### 2014-06-12 slides/2014_ics/slides-ics.tex
slieds

\only<1-2>{  \begin{block}{Shallow waters equations}
    \begin{eqnarray}
      \dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{ \textcolor{blue}{\partial}  u}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} u}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} u}{\textcolor{green}{\partial z}} + f(yw - zv) + g\dfrac{\textcolor{blue}{\partial} h}{\textcolor{blue}{\partial x}} \right) 
      \nonumber \\
      \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{ \textcolor{blue}{\partial} v}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} v}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} v}{\textcolor{green}{\partial z}} + f(zu - xw) + g\dfrac{\textcolor{red}{\partial} h}{\textcolor{red}{\partial} y} \right)
      \nonumber \\
      \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\textcolor{blue}{\partial} w}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} w}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} w}{\textcolor{green}{\partial z}} + f(xv - yu) + g\dfrac{\textcolor{green}{\partial} h}{\textcolor{green}{\partial z}} \right) 
      \nonumber\\
      \frac{\partial h}{\partial t} &=& -\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right)  \nonumber
    \end{eqnarray}
  \end{block}}

@@

  \begin{block}{Idea}
    \scriptsize
    \begin{equation}
      \left( \begin{array}{cccc}
          \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
          \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
          \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
          \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
        \end{array} \right)
      = \left(
        \begin{array}{c}
          \textcolor{blue}{D_{x}}\\ \textcolor{red}{D_{y}}\\ \textcolor{green}{D_{z}}\\ D_{hyp}
        \end{array}\right)
      \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) \nonumber
      \vspace{-1ex}
    \end{equation}
  \end{block}

@@

  \begin{block}{Why would it work}
    All the derivative matrices have the same sparsity pattern.

@@

  \begin{columns}
    \column{.4\linewidth}
    \scalebox{.55}{
      \begin{tabular}{|c|l|}
        % \hline
        % & & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
    \column{.6\linewidth}

@@

    Number of floating point operations:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$
  \end{columns}

  \begin{center}
    Intensity (Best case):\\
    $I_b = \frac{O}{b_T+ b_{wT}} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$

    Intensity (Worst case):\\
    $I_w = \frac{O}{B_T+ b_{wT}} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$
  \end{center}
\end{frame}

>>>>

\only<1-2>{  \begin{block}{Shallow waters equations}
    \vspace{-1em}
    \begin{eqnarray}
      \dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{ \textcolor{blue}{\partial}  u}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} u}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} u}{\textcolor{green}{\partial z}} + f(yw - zv) + g\dfrac{\textcolor{blue}{\partial} h}{\textcolor{blue}{\partial x}} \right) 
      \nonumber \\
      \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{ \textcolor{blue}{\partial} v}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} v}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} v}{\textcolor{green}{\partial z}} + f(zu - xw) + g\dfrac{\textcolor{red}{\partial} h}{\textcolor{red}{\partial} y} \right)
      \nonumber \\
      \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\textcolor{blue}{\partial} w}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} w}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} w}{\textcolor{green}{\partial z}} + f(xv - yu) + g\dfrac{\textcolor{green}{\partial} h}{\textcolor{green}{\partial z}} \right) 
      \nonumber\\
      \frac{\partial h}{\partial t} &=& -\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right)  \nonumber
    \end{eqnarray}
  \end{block}}

@@

  \begin{block}{Idea}
    \scriptsize
    \begin{equation}
      \left( \begin{array}{cccc}
          \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
          \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
          \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
          \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
        \end{array} \right)
      = \left(
        \begin{array}{c}
          \textcolor{blue}{D_{x}}\\ \textcolor{red}{D_{y}}\\ \textcolor{green}{D_{z}}\\ D_{hyp}
        \end{array}\right)
      \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) \nonumber
      \vspace{-1ex}
    \end{equation}
    \only<2>{$D_{hyp}$ is essentially a Laplacian operator used to adjust time stepping.}
  \end{block}

@@

  \begin{block}{Why would it work ?}
    All the derivative matrices have the same sparsity pattern.

@@

  \begin{columns}
    \column{.4\linewidth}
    \scalebox{.7}{
      \begin{tabular}{|c|l|}
        % \hline
        % & & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
%        $s_M$ & size of the $n_m$ matrices in bytes\\
%        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
%        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
%        \hline
        $cl$    & size of a cache line in bytes\\
%        $b_{wT}$ & number of bytes written to memory  \\
%        $b_{rT}$ & minimum number of bytes read from memory  \\
%        $b_T$   & minimum number of bytes transferred \\
%        $B_{rT}$ & maximum number of bytes read from memory \\
%        $B_T$   & maximum number of bytes transferred \\
%        $O$     & number of floating point operations \\
%        $I_b$   & maximum computational intensity\\
%        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }

    \includegraphics[width=\linewidth]{slides-figures/PPAM-figures/spmv.pdf}
    \column{.6\linewidth}

@@

    Number of floating point operations:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$

    Intensity (Best case):\\
%    $I_b = \frac{O}{b_T+ b_{wT}} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$
    $I_b = \frac{O}{b_T+ b_{wT}}$

    Intensity (Worst case):\\
%    $I_w = \frac{O}{B_T+ b_{wT}} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$
    $I_w = \frac{O}{B_T+ b_{wT}}$
  \end{columns}
\end{frame}
